Learning constraints in spreadsheets and tabular data

### 2016-09-24 Paper/ECAI/tabular.tex
slight retouch and some more comments by Tias

\tias{I'm up to this point}

@@

\paragraph{Q1: Effectiveness}
Our first questions, Q1, is about the inability to find intended constraints due to excessive runtime. As Table~\ref{tbl:category_overview} (last row) shows, the runtime of our method on these spreadsheets is always fast.  
The average runtime across all spreadsheets is $0.08s$.
Across categories the average running times fluctuate in the range between $0.04s$ to $0.20s$.
While most spreadsheets can be processed very quickly, running times can go up to around $0.65s$ for others.
In the previous section we analyzed the effect of various factors that can help explain runtime behavior.
Our benchmark experiments show that our algorithm is able to perform well in practice, even for larger spreadsheets.

@@

Our method uses dependencies to increase efficiency by finding constraints incrementally. To illustrate the effect this can have, we ran the benchmarks without using \textit{ORDERED} as base constraint for fuzzy lookup.
The total running times per category increase by 74\% (exercises), 84\% (tutorials) and~13\% (data).
This shows that the use of dependencies can have a strong effect on running times by sharing the computation and pruning done for one constraint with another.
 % (0.83 + 1.03 + 0.99) = 2.85 / (1.34 + 1.46 + 0.46) = 3.3  / (0.99 + 0.98 + 0.44) = 2.41

@@

\paragraph{Q2: Intended and supported constraints}
%\item Q2: it may fail to find intended constraints because it does not support such constraints; how does or method perform on real spreadsheets?
The overall recall achieved by our system, \sname, is~$0.88~(81/92)$.
When looking at the recall per category, \sname is able to consistently obtain high values for both overall and average recall, as shown in Table~\ref{tbl:category_overview}.
The intended constraints not discovered correspond to constraints that are not supported by the system (supported recall is 1).
Therefore, \sname is able to find all constraints for supported constraint templates, which confirms that our bias (signature) is not too restrictive.

@@

The intended constraints that are not discovered correspond with nested constraints (e.g. $R = \frac{W}{(L/100)^2}$), variants of supported constraints (e.g. series that do not start at 1) or more complex versions of supported constraints (e.g. conditional aggregates that automatically infer the condition).

\paragraph{Q3: Precision}
%\item Q3: it may find non-intended constraints; how many and what type of non-intended constraints are found on real spreadsheets?
The final question, Q3, is about the amount and type of non-intended constraints found.
Across all spreadsheets \sname achieves a precision of~$0.79~(81/103)$.
While the average precision per spreadsheet is high, over $0.90$ in every category, however the overall precision is much higher for the exercises and data categories ($0.94$ and $1.00$) than for the tutorials category ($0.70$).

Examining the tutorial spreadsheets confirms that there are a few spreadsheets that have a high number of additional constraints, e.g. an spreadsheet computes aggregates on its inventory but copies the data column for every aggregate.
Since every aggregate can be calculated based on any of the (equal) columns, many of the resulting constraints will be marked as unintended.
Usually, unintended constraints are the result of multiple ways to calculate the same result, e.g. if there is duplicate data.
A way to increase the precision is by post-processing the constraints to detect implications and heuristically selecting one among multiple ways to compute the same vector.

@@

\section{Applications}\label{sec:applications}
In this section, we illustrate how various motivating applications could be realized using our system.

@@

Consider Figure \ref{fig:autocompletion_example} containing two tables.
Learning on a snapshot that does not contain the row below $T_2$, we find constraints $\ecsumif{\range{T_2}{\rangeall}{2}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{3}}$ and $\ecsumif{\range{T_2}{\rangeall}{3}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{4}}$.
The new data $\Delta D$ consists of the new value \textit{West} in the row below $T_2$.
Therefore, vector~$\range{T_2}{:}{1}$ is changed and vectors~$\range{T_2}{:}{2}$, $\range{T_2}{:}{3}$ have become incomplete.
The \textit{SUMIF} constraints have enough input to fill the incomplete vectors and predict the values in the last row: $\range{T_2}{4}{2} = 1547$ and $\range{T_2}{4}{3} = 428128$.

@@

We consider two cases of error detection.
The first setting is the \textit{online} detection of errors and is similar to the autocompletion setting.
Using snapshots, we search for \textit{conflict} constraints, which are learned on a previous snapshot but do not agree with newly added data~$\Delta D$.
Such conflicts can indicate (typing) errors in the data just entered. For example, in case 'Wes' instead of 'West' was added in Figure~\ref{fig:autocompletion_example}.
%Conflict constraints can indicate errors in the input.
%However, such conflicts can be wrongly caused by spurious constraints.
%Hence, conflicts should perhaps only be detected if the snapshot data is large enough.
%An exact specification of what large enough means (i.e., how to measure the confidence in the constraint violation) is a separate application-dependent research question and will be investigated in the future work.

The second setting is \textit{offline} error detection in a static spreadsheet.
% Consider a use case concerning crime statistics provided by the FBI in Figure~\ref{fig:fbi}.
% Our system is able to detect the second sum constraint, however, a missing value prevents the first sum constraint from being learned.
% A possible approach to detect such missing or wrong values is to look at a sample of the table, e.g. trying to leave out each row, and compare the constraints learned on this sample with the constraints learned on the whole table.
  %The sample needs to be large enough such that we have enough confidence that the constraints are correct. For example, in Figure \ref{fig:fbi} all but one row satisfy the constraint, which is a strong indication that there is a constraint violation.
Instead of snapshots, one can repeatedly search for conflict constraints between the original spreadsheet and a sheet with a row or column removed. Many schemes can be devised for choosing which row(s) or column(s) are removed.
%This task can be formulated as online error detection by repeatedly partitioning the spreadsheet into a snapshot~$D_i$ and ``input'' $\Delta D_i$ (e.g. leave out each row and consider it as $\Delta D_i$) to detect conflicts.

@@

% points: 1) can be a base for an actual plugin or a function in excel 2) novel problem and challenge for systems and constraint solvers 3) can be part of the complicated pipeline togehter with the header detection
Our goal is to automatically identify constraints in a spreadsheet.
We have presented and evaluated our approach, implemented as the system \sname, that is able to learn many different constraints in tabular data.
The resulting method has high recall, produces limited number of redundant constraints and is sufficiently efficient for normal and interactive use.

>>>>

@@

\paragraph{Q1: Effectiveness}
Our first questions, Q1, is about the inability to find intended constraints due to excessive runtime. As Table~\ref{tbl:category_overview} (last row) shows, the runtime of our method on these spreadsheets is always fast.  
The average runtime across all spreadsheets is $0.08s$.
Across categories the average running times fluctuate in the range between $0.04s$ to $0.20s$.
While most spreadsheets can be processed very quickly, running times can go up to around $0.65s$ for others.
In the previous section we analyzed the effect of various factors that can help explain runtime behavior. \tias{Add something like 'on data generated to trigger many checks (equal sizes and lengths of many vectors)?? You choose}
Our benchmark experiments show that our algorithm is able to perform well in practice, even for larger spreadsheets.

@@

Our method uses dependencies to increase efficiency by finding constraints incrementally. To illustrate the effect this can have, we ran the benchmarks without using \textit{ORDERED} as base constraint for fuzzy lookup.
The total running times per category increase by 74\% (exercises), 84\% (tutorials) and~13\% (data).
This shows that the use of dependencies can have a strong effect on running times by sharing the computation and pruning done for one constraint with another.
 % (0.83 + 1.03 + 0.99) = 2.85 / (1.34 + 1.46 + 0.46) = 3.3  / (0.99 + 0.98 + 0.44) = 2.41
\tias{Very nice!!!}

@@

\paragraph{Q2: Intended and supported constraints}
%\item Q2: it may fail to find intended constraints because it does not support such constraints; how does or method perform on real spreadsheets?
The overall recall achieved by our system, \sname, is~$0.88~(81/92)$.
When looking at the recall per category, \sname is able to consistently obtain high values for both overall and average recall, as shown in Table~\ref{tbl:category_overview}.
The intended constraints not discovered correspond to constraints that are not supported by the system (supported recall is 1).
Therefore, \sname is able to find all constraints for supported constraint templates, which confirms that our bias (signature) is not too restrictive. \tias{'confirms'? We never write anywhere that we doubt this (and if we did I probably removed it : ///) Alternative could be something like "shows that our method is able to learn all constraints using the bias chosen (the formalism of signature and definition)"?}

@@

The intended constraints that are not discovered correspond with nested constraints (e.g. $R = \frac{W}{(L/100)^2}$), variants of supported constraints (e.g. series that do not start at 1) or more complex versions of supported constraints (e.g. conditional aggregates that automatically infer the condition). \tias{if you can clarify the latter, please do; if too difficult OK but 'automatically infer' sounds.. unusual for Excel}

\paragraph{Q3: Precision}
%\item Q3: it may find non-intended constraints; how many and what type of non-intended constraints are found on real spreadsheets?
The final question, Q3, is about the amount and type of \textbf{non}-intended constraints found.
Across all spreadsheets \sname achieves a precision of~$0.79~(81/103)$.
While the average precision per spreadsheet is high, over $0.90$ in every category, we observe that the overall precision is much higher for the exercises and data categories ($0.94$ and $1.00$) than for the tutorials category ($0.70$).

Examining the tutorial spreadsheets confirms that there are a few spreadsheets that have a high number of additional constraints, e.g. one spreadsheet computes aggregates on inventory data, but copies the data column for every aggregate.
Since every aggregate can be calculated based on any of the (equal) columns, many of the resulting constraints will be marked as unintended. \tias{rewrite 'many of the..' into 'our method will find all of these; only one in 'true' set, others drive down precision...'}

Such unintended constraints are the result of multiple ways to calculate the same result, e.g. if there is duplicate data.
One way to increase precision for this case is to post-process the constraints to detect equivalences and implications, and heuristically select one among multiple ways to compute the same vector.

@@

\section{Applications}\label{sec:applications}
In this section, we illustrate how various applications can build on our method for solving the tabular constraint learning problem.

@@

Consider Figure \ref{fig:autocompletion_example} containing two tables.
Learning on a snapshot that does not contain the row below $T_2$, our method finds constraints $\ecsumif{\range{T_2}{\rangeall}{2}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{3}}$ and $\ecsumif{\range{T_2}{\rangeall}{3}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{4}}$.
The new data $\Delta D$ consists of the new value \textit{West} in the row below $T_2$.
Therefore, vector~$\range{T_2}{:}{1}$ is changed and vectors~$\range{T_2}{:}{2}$, $\range{T_2}{:}{3}$ have become incomplete.
The \textit{SUMIF} constraints have enough input to fill the incomplete vectors and predict the values in the last row: $\range{T_2}{4}{2} = 1547$ and $\range{T_2}{4}{3} = 428128$.

@@

We consider two cases of error detection.
The first setting is the \textit{online} detection of errors and is similar to the autocompletion setting.
Using snapshots, one can learn the constraints in the previous snapshot and search for \textit{conflict} constraints, which do not agree with newly added data~$\Delta D$.
Such conflicts can indicate (typing) errors in the data just entered. For example, in case 'Wes' instead of 'West' was added in Figure~\ref{fig:autocompletion_example}.
%Conflict constraints can indicate errors in the input.
%However, such conflicts can be wrongly caused by spurious constraints.
%Hence, conflicts should perhaps only be detected if the snapshot data is large enough.
%An exact specification of what large enough means (i.e., how to measure the confidence in the constraint violation) is a separate application-dependent research question and will be investigated in the future work.

The second setting is \textit{offline} error detection in a static spreadsheet.
% Consider a use case concerning crime statistics provided by the FBI in Figure~\ref{fig:fbi}.
% Our system is able to detect the second sum constraint, however, a missing value prevents the first sum constraint from being learned.
% A possible approach to detect such missing or wrong values is to look at a sample of the table, e.g. trying to leave out each row, and compare the constraints learned on this sample with the constraints learned on the whole table.
  %The sample needs to be large enough such that we have enough confidence that the constraints are correct. For example, in Figure \ref{fig:fbi} all but one row satisfy the constraint, which is a strong indication that there is a constraint violation.
Instead of snapshots, one can repeatedly search for conflict constraints between the original spreadsheet and a sheet with a row or column removed. Many schemes can be devised for choosing which row(s) or column(s) are removed. \tias{Expand 'one can repeatedly search' into the actual steps: take spreadsheet, remove row/column, detect constraints and then search for any constraints not present in the full data (can easily be optimized to avoid known constraints btw)}
%This task can be formulated as online error detection by repeatedly partitioning the spreadsheet into a snapshot~$D_i$ and ``input'' $\Delta D_i$ (e.g. leave out each row and consider it as $\Delta D_i$) to detect conflicts.

@@

% points: 1) can be a base for an actual plugin or a function in excel 2) novel problem and challenge for systems and constraint solvers 3) can be part of the complicated pipeline togehter with the header detection
Our goal is to automatically identify constraints in a spreadsheet.
We have presented and evaluated our approach, implemented as the system \sname, that is able to learn many different constraints in tabular data.
The resulting method has high recall, produces a limited number of redundant constraints and is sufficiently efficient for normal and interactive use.
\tias{Smth about extendible: more constraints can be added by specifying signature in terms of properties and implementing the subblock finding (or in less detail, can be added, e.g. all native Excel constraints... should this be true)}
